---
title: "ggplot2"
author: "Alice Tivarovsky"
date: "9/26/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r df_1, cache = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```


#Start with an old plot 

```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5)

```

add labels: 

```{r}

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) + 
  labs(
    title = "temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)", 
    caption = "Data from NOAA via rnoaa package"
  )
```

x-axis tick marks ect

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) + 
  labs(
    title = "temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)", 
    caption = "Data from NOAA via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5C", "20C")
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
```


## Colors

```{r}
ggp_base = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.5) + 
  labs(
    title = "temperature plot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)", 
    caption = "Data from NOAA via rnoaa package"
  ) + 
  viridis::scale_color_viridis(name = "location", discrete = TRUE)

ggp_base
```

## Themes
themes are just things that don't fit into aesthetics/geoms; theme_bw is nice, theme_minimal is also nice; you can get rid of the legend by using legend.position = "none"; overarching themes go first, tweaks go next

```{r}
ggp_base + 
  theme_minimal() +
  theme(legend.position = "bottom")
```

You can also preset ggplot options in the beginning of your .rmd

You can also put two datasets in one plot using two geoms

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

# Patchwork

You can juxtapose several different plots using patchwork. First download the package from github

```{r}
devtools::install_github("thomasp85/patchwork")
```

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

# Data Manipulation

R uses alphabetial ordering, so Central Park is first; you can change this but need to make it a factor variable first, then use fct_relevel
```{r}
weather_df %>%
  mutate(name = factor(name), 
         name = fct_relevel(name, "Waikiki_HA", "CentralPark_NY")) %>% 
ggplot(aes(x = name, y = tmax, color = name)) + geom_boxplot()
  
```

